%%
\section{Conclusion}

%% Cat-and-mouse game
Decades after its introduction, web tracking still remains an archetypal cat-and-mouse game.
%
Each incremental defense---whether a new browser policy or a regulatory ruling---quickly provokes an equally sophisticated evasion technique to track users.
%
This adversarial dynamic shows that purely reactive approaches cannot deliver privacy guarantees for online users.

%%
Regulations alone are insufficient -- data protection statutes such as GDPR and CCPA have tightened accountability, yet such enforcement lags the speed of technical changes in evolving tracking mechanisms.
%
Moreover, trackers often find tolerated gray zones to bypass regulations.
%
As a result, enforcement frequently stalls on jurisdictional or interpretative disputes. 
%
There is a need for regulators to incorporate agile, evidence-driven auditing methods by collaborating with the measurement community to avoid any oversight and to ensure that regulations evolve competitively with the technical reality.

%% Browsers can be unreliable
On the other hand, while browsers are powerful gatekeepers, they provide an unreliable line of defense.
%
Default protections vary widely across browser vendors, experimental features sometimes ship years after the issues are identified, and commercial incentives often result in more permissive designs. 
%
Future research must therefore look beyond ``\textit{fix it in the browser}'' remedies and explore complementary approaches that truly safeguard user's privacy.

%% give control to users who want it
Thus, while browser-based protections and policy-driven changes are effective to some extent, current tracking landscape demands a default \textit{privacy-first} solution where users can control their privacy as opposed to browsers or regulators.
%
This SoK highlights this by summarizing important findings in the evolution of web tracking and its prevention across the years and suggesting key future directions.
%
Our hope is to aid informing the research community on what is novel and important to focus on in order to improve the state of user's online privacy.
